R Notebook

### 2022-04-02 Project/ltrs data.Rmd
Update ltrs data.Rmd

# Load data
ltrsHB <- read.csv("./HargreavesEcolLett data ltrsHB.csv")
ltrsNK <- read.csv("./HargreavesEcolLett data ltrsNK.csv")

@@

@@

@@

### Create Distance Matrix

# Select plant ID and physical parameters to compare similarity and create unique plant ID
ltrs_OTU_ID <- ltrsdat %>% 
    group_by(transect, Source, plant) %>%
    mutate(dupID = row_number()) %>%
    ungroup() %>% 
    mutate(plantID = paste(transect, Source, plant, dupID), sep = "_") %>% 
    select(plantID, tbo:totbadseed)

# Replace column numbers with plant ID
ltrs_OTU <- data.frame(ltrs_OTU_ID[,-1], row.names = ltrs_OTU_ID$plantID)

# Build distance matrix
# Create distance values
ltrs_Dist <- dist(ltrs_OTU, method = 'euclidean')

# Reshape as matrix
ltrs_DM <- as.matrix(ltrs_Dist)

# Reshape to be plotted
ltrs_plotdat <- melt(ltrs_DM)

# Distance matrix plot
ltrsDistMatrix <- ggplot(ltrs_plotdat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = c("white", "blue", "green", "red")) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())

ltrsDistMatrix

@@

### Visualize similarities with Neighbour-joining tree

```{r}
# NJ method
ltrs_tree <- nj(ltrs_Dist)

# Group by Source
SourceList <- gsub("\\w+\\s+(\\w).+","\\1", ltrs_tree$tip.label)
SourceGroup <- split(ltrs_tree$tip.label, SourceList)
SourceTree <- groupOTU(ltrs_tree, SourceGroup)

# Create tree
ltrsNJTree <- ggtree(SourceTree, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsNJTree

```
### Visualise dissimilarities with Bray-Cirtis Dissimilarity
```{r}

# Create distance matrix and tree
ltrs_bray <- vegdist(ltrs_OTU, method = "bray", binary = F, na.rm = T)
ltrs_tree_bray <- nj(ltrs_bray)

# Group by Source
SourceGroupBray <- split(ltrs_tree_bray$tip.label, SourceList)
SourceTreeBray <- groupOTU(ltrs_tree_bray, SourceGroupBray)

# Create tree
ltrsBrayTree <- ggtree(SourceTreeBray, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsBrayTree

@@

### Examine Data with Non-metric Multidimensional Scaling NA REMOVED

@@

# Remove NAs from dataset
ltrs_OTU_NA <- na.omit(ltrs_OTU)

# Convert dataframe to matrix
character_matrix <- as.matrix(ltrs_OTU_NA)

# Perform NMDS on community data
character_NMDS <-  metaMDS(character_matrix, k = 2, trymax = 100)

@@

### Examine Data with Non-metric Multidimensional Scaling BINARY ENCODE

# Check which columns have na values
colnames(ltrs_OTU)[colSums(is.na(ltrs_OTU)) > 0]

# Convert aveseed to binary
ltrs_bin <- ltrs_OTU %>% replace(is.na(.), 0)
ltrs_bin$aveseed1 = ifelse(ltrs_bin$aveseed1 > 1, 1, 0)
ltrs_bin$aveseed2 = ifelse(ltrs_bin$aveseed2 > 1, 1, 0)

# Convert dataframe to matrix
character_matrix <- as.matrix(ltrs_bin)

# Perform NMDS on community data
character_NMDS <-  metaMDS(character_matrix, k = 2, trymax = 500)

>>>>

setwd("~/GitHub/432_Group5/Project")

# Load data
ltrsHB <- read.csv("./Data/HargreavesEcolLett data ltrsHB.csv")
ltrsNK <- read.csv("./Data/HargreavesEcolLett data ltrsNK.csv")

@@

# Remove data not from High, Medium, or Low elevation 
elevation <- c("H", "M", "L")
ltrsdat <- ltrsdat %>% 
  filter(Source %in% elevation)# %>% 
  #mutate(site = recode(site, A1 = "A")) %>% # Merge site A1 and A2 data into A to simplify
  #mutate(site = recode(site, A2 = "A")) %>% 
  #mutate(site = recode(site, HE = "H")) # Merge site HE into H to simplify

@@

# Subset into High, Medium, and Low Source Elevation
ltrsHigh <- ltrsdat %>% 
    filter(Source == "H") %>% 
    group_by(transect, site, plant) %>%
    mutate(dupID = row_number()) %>%
    ungroup() %>% 
    mutate(plantID = paste(transect, site, plant, dupID), sep = "_") %>% 
    select(plantID, tbo:totbadseed)

ltrsMed <- ltrsdat %>% 
    filter(Source == "M") %>% 
    group_by(transect, site, plant) %>%
    mutate(dupID = row_number()) %>%
    ungroup() %>% 
    mutate(plantID = paste(transect, site, plant, dupID), sep = "_") %>% 
    select(plantID, tbo:totbadseed)

ltrsLow <- ltrsdat %>% 
    filter(Source == "L") %>% 
    group_by(transect, site, plant) %>%
    mutate(dupID = row_number()) %>%
    ungroup() %>% 
    mutate(plantID = paste(transect, site, plant, dupID), sep = "_") %>% 
    select(plantID, tbo:totbadseed)

# Replace column numbers with plantID
ltrsH_OTU <- data.frame(ltrsHigh[,-1], row.names = ltrsHigh$plantID)
ltrsM_OTU <- data.frame(ltrsMed[,-1], row.names = ltrsMed$plantID)
ltrsL_OTU <- data.frame(ltrsLow[,-1], row.names = ltrsLow$plantID)

@@

### Examine Dissimilarities for High Elevation Sourced Plants

Create Distance Matrix
```{r}

# Create distance values
ltrsH_Dist <- dist(ltrsH_OTU, method = 'euclidean')

# Reshape as matrix
ltrsH_DM <- as.matrix(ltrsH_Dist)

# Reshape to be plotted
ltrsH_plotdat <- melt(ltrsH_DM)

# Distance matrix plot
HDM_Plot <- ggplot(ltrsH_plotdat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = c("white", "blue", "green", "red")) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())

HDM_Plot

@@

Visualise similarities with Neighbour-Joining tree
```{r}
# NJ method
ltrsH_tree <- nj(ltrsH_Dist)

# Group by Source
SourceListH <- gsub("\\w+\\s+(\\w+)\\s.*","\\1", ltrsH_tree$tip.label)
SourceGroupH <- split(ltrsH_tree$tip.label, SourceListH)
SourceTreeH <- groupOTU(ltrsH_tree, SourceGroupH)

# Create tree
ltrsHNJTree <- ggtree(SourceTreeH, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsHNJTree

@@

Visualise dissimilarities with Bray-Curtis Dissimilarity
```{r}

# Create distance matrix and tree
ltrsH_bray <- vegdist(ltrsH_OTU, method = "bray", binary = F, na.rm = T)
ltrsHTree_bray <- nj(ltrsH_bray)

# Group by Source
SourceGroupHBray <- split(ltrsHTree_bray$tip.label, SourceListH)
SourceTreeHBray <- groupOTU(ltrsHTree_bray, SourceGroupHBray)

# Create tree
ltrsHBrayTree <- ggtree(SourceTreeHBray, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsHBrayTree

@@

# Check which columns have na values
colnames(ltrsH_OTU)[colSums(is.na(ltrsH_OTU)) > 0]

# Convert aveseed to binary
ltrsH_bin <- ltrsH_OTU %>% replace(is.na(.), 0)
ltrsH_bin$aveseed1 = ifelse(ltrsH_bin$aveseed1 > 1, 1, 0)
ltrsH_bin$aveseed2 = ifelse(ltrsH_bin$aveseed2 > 1, 1, 0)

# Convert dataframe to matrix
character_matrix_H <- as.matrix(ltrsH_bin)

# Perform NMDS on community data
character_NMDS_H <-  metaMDS(character_matrix_H, k = 2, trymax = 500)

### Examine Dissimilarities for Medium Elevation Sourced Plants

Create Distance Matrix
```{r}

# Create distance values
ltrsM_Dist <- dist(ltrsM_OTU, method = 'euclidean')

# Reshape as matrix
ltrsM_DM <- as.matrix(ltrsM_Dist)

# Reshape to be plotted
ltrsM_plotdat <- melt(ltrsM_DM)

# Distance matrix plot
MDM_Plot <- ggplot(ltrsM_plotdat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = c("white", "blue", "green", "red")) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())

MDM_Plot

Visualise similarities with Neighbour-Joining tree
```{r}
# NJ method
ltrsM_tree <- nj(ltrsM_Dist)

# Group by Source
SourceListM <- gsub("\\w+\\s+(\\w+)\\s.*","\\1", ltrsM_tree$tip.label)
SourceGroupM <- split(ltrsM_tree$tip.label, SourceListM)
SourceTreeM <- groupOTU(ltrsM_tree, SourceGroupM)

# Create tree
ltrsMNJTree <- ggtree(SourceTreeM, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsMNJTree

Visualise dissimilarities with Bray-Curtis Dissimilarity
```{r}

# Create distance matrix and tree
ltrsM_bray <- vegdist(ltrsM_OTU, method = "bray", binary = F, na.rm = T)
ltrsMTree_bray <- nj(ltrsM_bray)

# Group by Source
SourceGroupMBray <- split(ltrsMTree_bray$tip.label, SourceListM)
SourceTreeMBray <- groupOTU(ltrsMTree_bray, SourceGroupMBray)

# Create tree
ltrsMBrayTree <- ggtree(SourceTreeMBray, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsMBrayTree

# Check which columns have na values
colnames(ltrsM_OTU)[colSums(is.na(ltrsM_OTU)) > 0]

# Convert aveseed to binary
ltrsM_bin <- ltrsH_OTU %>% replace(is.na(.), 0)
ltrsM_bin$aveseed1 = ifelse(ltrsM_bin$aveseed1 > 1, 1, 0)
ltrsM_bin$aveseed2 = ifelse(ltrsM_bin$aveseed2 > 1, 1, 0)

# Convert dataframe to matrix
character_matrix_M <- as.matrix(ltrsM_bin)

# Perform NMDS on community data
character_NMDS_M <-  metaMDS(character_matrix_M, k = 2, trymax = 500)

### Examine Dissimilarities for Low Elevation Sourced Plants

Create Distance Matrix
```{r}

# Create distance values
ltrsL_Dist <- dist(ltrsL_OTU, method = 'euclidean')

# Reshape as matrix
ltrsL_DM <- as.matrix(ltrsL_Dist)

# Reshape to be plotted
ltrsL_plotdat <- melt(ltrsL_DM)

# Distance matrix plot
LDM_Plot <- ggplot(ltrsL_plotdat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = c("white", "blue", "green", "red")) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())

LDM_Plot

Visualise similarities with Neighbour-Joining tree
```{r}
# NJ method
ltrsL_tree <- nj(ltrsL_Dist)

# Group by Source
SourceListL <- gsub("\\w+\\s+(\\w+)\\s.*","\\1", ltrsL_tree$tip.label)
SourceGroupL <- split(ltrsL_tree$tip.label, SourceListL)
SourceTreeL <- groupOTU(ltrsL_tree, SourceGroupL)

# Create tree
ltrsLNJTree <- ggtree(SourceTreeL, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsLNJTree

Visualise dissimilarities with Bray-Curtis Dissimilarity
```{r}

# Create distance matrix and tree
ltrsL_bray <- vegdist(ltrsL_OTU, method = "bray", binary = F, na.rm = T)
ltrsLTree_bray <- nj(ltrsL_bray)

# Group by Source
SourceGroupLBray <- split(ltrsLTree_bray$tip.label, SourceListL)
SourceTreeLBray <- groupOTU(ltrsLTree_bray, SourceGroupLBray)

# Create tree
ltrsLBrayTree <- ggtree(SourceTreeLBray, layout = "circular", aes(colour = group)) +
  geom_tiplab(size = 2, aes(angle = 0)) +
  theme(legend.position = "right")

ltrsLBrayTree

# Check which columns have na values
colnames(ltrsL_OTU)[colSums(is.na(ltrsL_OTU)) > 0]

# Convert aveseed to binary
ltrsL_bin <- ltrsH_OTU %>% replace(is.na(.), 0)
ltrsL_bin$aveseed1 = ifelse(ltrsL_bin$aveseed1 > 1, 1, 0)
ltrsL_bin$aveseed2 = ifelse(ltrsL_bin$aveseed2 > 1, 1, 0)

# Convert dataframe to matrix
character_matrix_L <- as.matrix(ltrsL_bin)

# Perform NMDS on community data
character_NMDS_L <-  metaMDS(character_matrix_L, k = 2, trymax = 500)

@@
